### 2019-06-20 submission/manuscript.Rmd
Preliminary draft

## Introduction
Proton pump inhibitors (PPIs) are one of the most widely used drugs and are available over the counter. PPIs are approved to treat a variety of conditions including gastroesophageal reflux disease, ulcers, and prophylaxis for nonsteroidal anti-inflammatory drug (NSAID)-associated gastrointestinal bleeding [@nehra2018proton]. Unfortunately, there have been a number of adverse events associated with their use including small intestinal bacterial overgrowth, decreased calcium absorption, and increased risk of enteric infections [@nehra2018proton]. There’s also been a number of large cohort studies as well as interventional clinical trials that have demonstrated specific alterations in the intestinal microbiome associated with PPI use [@Imhann2017; @Naito2018]. PPI-associated microbiota changes have been attributed to the PPIs ability to increase stomach pH which may promote the survival of oral and pathogenic bacteria. Human fecal microbiota changes with PPI use include increases in Enterococcaceae, Lactobacillaceae, Micrococcaceae, Staphylococcaceae and Streptococcaceae with decreases in Ruminococcaceae [@Imhann2017]. An in vitro screen, demonstrated that 3 PPIs inhibit the growth of 7-8 strains of Lachnospiraceae and Ruminococcaceae, which represented ~16% of the 40 intestinal bacteria strains that were screened [@maier2018extensive].
Antibiotics have a large impact on the intestinal microbiome and are a primary risk factor for developing CDIs [@Schubert2015]. It is less clear whether other human medications that impact the microbiota also influence *C. difficile* colonization resistance. Multiple epidemiological studies have suggested an association between PPI use and incidence of Clostridium difficile infections (CDIs) [@tariq2017association; @nehra2018proton; @Naito2018]. However, most of the studies suggesting a link between PPIs and *C. difficile* are retrospective and did not evaluate the microbiome [@tariq2017association; @nehra2018proton]. Thus, it is unclear whether the gastrointestinal microbiome changes associated with PPI use play a role in the association between PPIs and CDI incidence, severity or recurrence. Additionally, epidemiological studies have a limited capacity to address potential confounders and comorbidities in patients that were on PPIs and developed CDI infections. 
Although there have been a few *C. difficile* mouse model studies that have demonstrated PPIs have some effect on *C. difficile* colonization with or without additional antibiotic treatment, none of these studies looked at the interplay with the intestinal microbiome [@hung2015proton; @Kaur2007; @kaur2011effect]. Here, we evaluated the impact of a PPI on the murine microbiome and susceptibility to C. difficile colonization in relation to clindamycin, an antibiotic that perturbs the microbiome enough to allow C. difficile to colonize but is mild enough that C. difficile is cleared within 10 days [@Jenior2018].

>>>>

## Introduction
Proton pump inhibitors (PPIs) are one of the most widely used drugs and are available over the counter. PPIs are approved to treat a variety of conditions including gastroesophageal reflux disease, ulcers, and prophylaxis for nonsteroidal anti-inflammatory drug (NSAID)-associated gastrointestinal bleeding [@nehra2018proton]. Unfortunately, there have been a number of adverse events associated with their use including small intestinal bacterial overgrowth, decreased calcium absorption, and increased risk of enteric infections [@nehra2018proton]. There’s also been a number of large cohort studies as well as interventional clinical trials that have demonstrated specific alterations in the intestinal microbiome associated with PPI use [@Imhann2017; @Naito2018]. PPI-associated microbiota changes have been attributed to the PPIs ability to increase stomach pH which may promote the survival of oral and pathogenic bacteria. Human fecal microbiota changes with PPI use include increases in Enterococcaceae, Lactobacillaceae, Micrococcaceae, Staphylococcaceae and Streptococcaceae with decreases in Ruminococcaceae [@Imhann2017]. An in vitro screen, demonstrated that 3 PPIs inhibit the growth of 7-8 strains of Lachnospiraceae and Ruminococcaceae, which represented ~16% of the 40 intestinal bacteria strains that were screened [@maier2018extensive].
Antibiotics have a large impact on the intestinal microbiome and are a primary risk factor for developing CDIs [@Schubert2015]. It is less clear whether other human medications that impact the microbiota also influence *C. difficile* colonization resistance. Multiple epidemiological studies have suggested an association between PPI use and incidence of Clostridium difficile infections (CDIs) [@tariq2017association; @nehra2018proton; @Naito2018; @Elias2019]. However, most of the studies suggesting a link between PPIs and *C. difficile* are retrospective and did not evaluate the microbiome [@tariq2017association; @nehra2018proton; @Elias2019]. Thus, it is unclear whether the gastrointestinal microbiome changes associated with PPI use play a role in the association between PPIs and CDI incidence, severity or recurrence. Additionally, epidemiological studies have a limited capacity to address potential confounders and comorbidities in patients that were on PPIs and developed CDI infections. 
Although there have been a few *C. difficile* mouse model studies that have demonstrated PPIs have some effect on *C. difficile* colonization with or without additional antibiotic treatment, none of these studies looked at the interplay with the intestinal microbiome [@hung2015proton; @Kaur2007; @kaur2011effect]. Here, we evaluated the impact of a PPI on the murine microbiome and susceptibility to C. difficile colonization in relation to clindamycin, an antibiotic that perturbs the microbiome enough to allow C. difficile to colonize but is mild enough that C. difficile is cleared within 10 days [@Jenior2018].
